---
title: "20241118_course_part1"
author: "camilsek"
date: "`r Sys.Date()`"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

# Course part 1 - laboratory

-   We selected 20 isolates for subculturing
    -   6 isolates are *Klebsiella* spp. isolates
    -   6 isolates are *Escherichia coli* isolates
    -   8 isolates are *Salmonella* spp. isolates

# Course part 2
- Installation of packages
```{r}
```

```{r}
#install.packages("here")
install.packages("DiagrammeR")
```
Only need to install a library once!!

# ctrl + alt + i will give the start and stop of the code (cell)

```{r}
angry_or_not <- "not"
```
gives content of the object
```{r}
angry_or_not
```


Madelaine, are you angry or not? `r print(angry_or_not)`

#Working on the human data


```{r}
here::here()
```


```{r}

path_data <- here::here("data/2024-09-25_INIKA_SAMPLING_ANALYSIS_HUMAN.csv")

path_data
```

```{r}
human_data <-
  readr::read_csv2(path_data)
```

day3
#Packagees into memory

```{r}
library (here)
library(tidyverse)
```

```{r}
?library()
help("library")
```

Another way of using help - se på "Help" fane bruk F1 i selve funksjonen feks funksjonen library(here) sett kursor på library før du trykker. 

#We get our data

```{r}

path_data <- here::here("data/2024-09-25_INIKA_SAMPLING_ANALYSIS_HUMAN.csv")
path_data
```

```{r}
human_data <-
  readr::read_csv2(path_data)
```
## Exploration of a data frame
- Name of the columns
```{r}
colnames(human_data)
```

- getting the column names into a vector for pasting and manipulation
```{r}
dput(colnames(human_data))
```
- explanation: c() is to make vectors

```{r}
test <- c("one", "2", 3)
test
typeof(test)
class(test)
```
- Vectors in R are all of the same type, if text called "character", usually chose and base on the first.
- Overview of data and data types


```{r}
test <- c("one", "2", 3)
test
test[3]
typeof(test[3])
typeof(test)
class(test)
```

```{r}
test2 <- c(1, 2, 3)
typeof(test2)
class(test2)
test2*test2
```
```{r}
test2 * c(1,2)
```

- deactivate the output in the markdown document settings "chunk output un the console".

Need to reduce number of columns in the human dataset

```{r}
glimpse(human_data)
str(human_data)
```

- renaming column names (easier to use)
```{r}
original_names <- colnames(human_data)
original_names
```

```{r}
new_names <- str_replace_all(original_names, " ","_")
new_names
new_names[88]
```

Remove all strange characters
- regex pattern
```{r}
new_names2 <- str_replace_all(new_names, "[?/,;.*()-]","_")
new_names2
new_names2[88]
```

```{r}
colnames(human_data) <- new_names2
```

```{r}
colnames(human_data) <- new_names2
colnames(human_data)[88]
```

- learn how to select columns and transfer the data set we need. Want to reduce size of the table.

```{r}
human_data%>%
  select(INIKA_OH_TZ_ID,Age__yrs_,Gender) %>%
  head()
```

```{r}
human_data_selection <- 
  human_data %>%
  select(INIKA_OH_TZ_ID, Age__yrs_, Gender, Enter_a_date, Region, District, 
         Specify_if_other_district, Sample, Season, Origin_of_sample, 
         Which_class_grade_are_you__, 
         Who_is_your_caretaker__, 
         If_others__mention, 
         What_is_your_occupation_and_or_of_your_caretaker__, 
         Have_you_ever_heard_about_AMR_, If_yes__how_did_you_get_this_information_, 
         Have_you_or_your_children_used_any_antibiotics_at_any_time_, 
         If_yes__where_did_you_get_these_drugs_from_,
         If_it_was_drug_sellers_or_pharmacy__did_you_have_a_prescription_from_the_doctor_prescriber_,
         GPS_coordinates) 

human_data_selection %>% View()
```

Gjorde endringer i skriptet manuelt, men det er en måte å fjerne både underscore foran og bak i fila: str_remove_all("(^_*)|(_*$)").

```{r}
human_data_selection %>% head()
```

```{r}
glimpse(human_data_selection)
```

Gives you the type, dbl, chr, etc

```{r}
str(human_data_selection)
```
str also gives you some information about the data in the different columns

remove column

```{r}
human_data_selection %>%
  select(-Specify_if_other_district) %>%
  mutate_at(vars(INIKA_OH_TZ_ID), as.character) %>%
  mutate_at(vars(Age__yrs_), as.integer) %>%
  str()
```

Difference between categorical variable and character?

```{r}
human_data_selection %>%
  select(-Specify_if_other_district) %>%
  mutate_at(vars(INIKA_OH_TZ_ID), as.character) %>%
  mutate_at(vars(Age__yrs_), as.integer) %>%
  mutate_if(is.character, factor) %>%
  str()
```

Endre fra character til factor

```{r}
human_data_selection <-
human_data_selection %>%
  select(-Specify_if_other_district) %>%
  mutate_at(vars(Age__yrs_), as.integer) %>%
  mutate_if(is.character, factor) %>%
  mutate_at(vars(INIKA_OH_TZ_ID), as.character)
```

When happy with the changes assign or reassign to the name by <-
Remember to remove %>%

```{r}
summary(human_data_selection)
```

```{r}
human_data_selection %>%
  filter(INIKA_OH_TZ_ID == "238")
```

```{r}
human_data_selection %>%
  filter(INIKA_OH_TZ_ID != "238") %>%
  filter(is.na(Gender))
```
