---
title: "20241118_course_part1"
author: "camilsek"
date: "`r Sys.Date()`"
output:

  html_document: default

---

# Course part 1 - laboratory

-   We selected 20 isolates for subculturing
    -   6 isolates are *Klebsiella* spp. isolates
    -   6 isolates are *Escherichia coli* isolates
    -   8 isolates are *Salmonella* spp. isolates

# Course part 2
- Installation of packages
```{r}
```

```{r}
#install.packages("here")
install.packages("DiagrammeR")
```
Only need to install a library once!!

# ctrl + alt + i will give the start and stop of the code (cell)

```{r}
angry_or_not <- "not"
```
gives content of the object
```{r}
angry_or_not
```


Madelaine, are you angry or not? `r print(angry_or_not)`

#Working on the human data


```{r}
here::here()
```


```{r}

path_data <- here::here("data/2024-09-25_INIKA_SAMPLING_ANALYSIS_HUMAN.csv")

path_data
```

```{r}
human_data <-
  readr::read_csv2(path_data)
```

day3
#Packagees into memory

```{r}
library (here)
library(tidyverse)
```

```{r}
?library()
help("library")
```

Another way of using help - se på "Help" fane bruk F1 i selve funksjonen feks funksjonen library(here) sett kursor på library før du trykker. 

#We get our data

```{r}

path_data <- here::here("data/2024-09-25_INIKA_SAMPLING_ANALYSIS_HUMAN.csv")
path_data
```

```{r}
human_data <-
  readr::read_csv2(path_data)
```
## Exploration of a data frame
- Name of the columns
```{r}
colnames(human_data)
```

- getting the column names into a vector for pasting and manipulation
```{r}
dput(colnames(human_data))
```
- explanation: c() is to make vectors

```{r}
test <- c("one", "2", 3)
test
typeof(test)
class(test)
```
- Vectors in R are all of the same type, if text called "character", usually chose and base on the first.
- Overview of data and data types


```{r}
test <- c("one", "2", 3)
test
test[3]
typeof(test[3])
typeof(test)
class(test)
```

```{r}
test2 <- c(1, 2, 3)
typeof(test2)
class(test2)
test2*test2
```
```{r}
test2 * c(1,2)
```

